Learning words by recursive pragmatic reasoning about other agents

### 2013-05-31 praglearn.tex
edits to model section

\subsection{Paradoxes in optimal models of pragmatic learning}
% When modelling human behavior, it's useful to begin with optimal
% models; this has the advantage that when we are then forced to
% deviate from the purely rational approach, it's generally in a
% theoretically interpretable way. However,
Multi-agent interactions are difficult to model in a normative or
optimal framework without falling prey to paradox. For example,
consider a simple model of the agents in the above game. First we
define a \textit{literal listener} $L_0$. This agent has a
\textit{lexicon} which probabilistically maps words to meanings;
specifically, it assigns each word $w$ a vector describing the extent
to which this word provides evidence for each object\footnote{We
  assume words refer directly to objects, rather than to abstract
  semantic features. Our simplification is without loss of
  generalization, however, because we can interpret our model as
  marginalizing over such a representation, with our literal
  $P_\lex(\obj | \word) = \sum_\text{features} P(\obj |
  \text{features}) P_\lex(\text{features} | \word)$.}. Each entry in
the vector is in $(0, 1)$, and together they sum to 1:
\begin{align*}
P_{L_0}(\obj | \word, \lex) &\propto \lex(\word, \obj) \times \prior.
\end{align*}
Confronted with such a listener, a probabilistically rational speaker
should attempt to choose a word which soft-maximizes the probability
that the listener will assign to the target object---modulated by the
effort or cost associated with producing this word:
\begin{align*}
P_{S_1}(\word | \obj, \lex) &\propto \exp\Big(\lambda \big(\log P_{L_0}(\obj | \word, \lex) - \text{cost}(\word)\big)\Big).
\end{align*}
But given this speaker, then the naive $L_0$ strategy is not
optimal. Instead, listeners should use Bayes rule to invert the
speaker's decision procedure \cite{frank2012}:
\begin{align*}
P_{L_2}(\obj | \word, \lex) &\propto P_{S_1}(\word | \obj, \lex) \times \prior.
\end{align*}
Now the paradox becomes apparent. Given such a listener, it is no
longer optimal for speakers to implement strategy $S_1$; instead, they
should implement strategy $S_3$ which soft-maximizes $P_{L_2}$ instead
of $P_{L_0}$. And then listeners ought to implement $L_4$, and so on.

@@

If each agent in a multi-agent setting assumes the existence of a
shared, conventional norm underlying everyone's behaviour, and further
assumes that other agents are knowledgeable about this norm while they
themselves are not, then we are able to preserve most of the normative
Bayesian framework, explicitly model both learning and recursive
social reasoning simultaneously within the same framework, and avoid
paradox. The resulting theory unifies two previously separate
strands of modelling work on language usage and acquisition,
accounting for a superset of previous phenomena, while producing a
variety of new qualitative patterns. In particular, we make new

>>>>

\subsection{Paradoxes in optimal models of pragmatic learning}
% When modelling human behavior, it's useful to begin with optimal
% models; this has the advantage that when we are then forced to
% deviate from the purely rational approach, it's generally in a
% theoretically interpretable way. However,
Multi-agent interactions are difficult to model in a normative or
optimal framework without falling prey to paradox. For example,
consider a simple model of the agents in the above game. First we
define a \textit{literal listener} $L_0$. This agent has a
\textit{lexicon} which probabilistically maps words to meanings;
specifically, it assigns each word $w$ a vector describing the extent
to which this word provides evidence for each possible object\footnote{We
  assume words refer directly to objects, rather than to abstract
  semantic features. Our simplification is without loss of
  generalization, however, because we can interpret our model as
  marginalizing over such a representation, with our literal
  $P_\lex(\obj | \word) = \sum_\text{features} P(\obj |
  \text{features}) P_\lex(\text{features} | \word)$.}. Each entry in
the vector is in $(0, 1)$, and together they sum to 1. By Bayes rule:
\begin{align*}
P_{L_0}(\obj | \word, \lex) &\propto \lex(\word, \obj) \times \prior.
\end{align*}
Confronted with such a listener, a probabilistically rational speaker
should attempt to choose a word which soft-maximizes the probability
that the listener will assign to the target object---modulated by the
effort or cost associated with producing this word:
\begin{align*}
P_{S_1}(\word | \obj, \lex) &\propto \exp\Big(\lambda \big(\log P_{L_0}(\obj | \word, \lex) - \text{cost}(\word)\big)\Big).
\end{align*}
But given this speaker, then the naive $L_0$ strategy is not
optimal. Instead, listeners should use Bayes rule to invert the
speaker's decision procedure \cite{frank2012}:
\begin{align*}
P_{L_2}(\obj | \word, \lex) &\propto P_{S_1}(\word | \obj, \lex) \times \prior.
\end{align*}
Now the paradox becomes apparent. Given such a listener, it is no
longer optimal for speakers to implement strategy $S_1$; instead, they
should implement strategy $S_3$ which soft-maximizes $P_{L_2}$ instead
of $P_{L_0}$. And then listeners ought to implement $L_4$, and so on.

@@

allows us to preserve most of the normative Bayesian framework within
the multi-agent setting, while avoiding paradoxes and allowing both
pragmatic reasoning and learning. The resulting model not only unifies
previous models of each phenomenon separately, it also accounts for a
range of
